---
title: "hw03_Ninadh"
author: "Ninadh"
date: "October 15, 2018"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(gapminder)
library(knitr)
```

```{r}
gapminder
```

task 1: maximum and minimum GDP per capita for all continents

```{r}
gapminder %>%
  group_by(continent) %>%
  summarise(minimum = min(gdpPercap),
            maximum = max(gdpPercap))
```

```{r}
# figure

gapminder %>%
  group_by(continent) %>%
  ggplot(aes(continent, gdpPercap, fill= continent)) + geom_violin()
```

Task 2: GDP spread within the continents

```{r}
gapminder %>%
  group_by(continent) %>%
  summarise(gdpSpread = max(gdpPercap) - min(gdpPercap))
```

Task 3: Life expectancy changing over time in different continents 

| Grammar Component     | Specification |
|-----------------------|---------------|
| __data__              | `gapminder` |
| __aesthetic mapping__ | 'x' & 'y' |
| __geometric object__  | point |
| scale                 | linear |
| statistical transform | none |

```{r}
gapminder %>%
  group_by(continent) %>%
  ggplot(aes(year, lifeExp)) +
  geom_point(aes(colour = continent)) +
  geom_smooth(aes(colour = continent))
```

Life expectancy of Oceania > Europe > Americas > Asia > Africa

Task 4: Continents with the number of people who had life expectancy below the mean life expectancy.

```{r}
gapminder %>%
  group_by(year, continent) %>%
  summarise(mean_lifeExp = mean(lifeExp), 
            low_lifeExp = sum(lifeExp < mean_lifeExp))

```


